\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-09-16 NAR_DRAFT.tex
add observation about APOBEC signature in 1F

\begin{figure}
  \begin{center}
  \includegraphics{figures/signatures_only.pdf}
  \end{center}
  \caption{Tumor mutation profiles.
  In each panel, the y-axis is the proportion of total mutations in the tumor with each of the 96 substitution types shown on the x-axis.
  \textbf{(A)} The observed mutation profile of an acute myeloid leukemia used in our real data analysis \cite{Griffith2015}.
  \textbf{(B)} The observed mutation profile of a breast tumor we used in our real data analysis \cite{Shi2018}.
  \textbf{(C)} A mutation profile used for simulating tumors, made up of equal proportions of COSMIC mutation signatures 1, 7, \& 11.
  \textbf{(D)} Equal proportions of signatures 1, 4, \& 5.
  \textbf{(E)} Equal proportions of signatures 1, 3, \& 5.
  \textbf{(F)}--\textbf{(G)} The observed mutation profile of two real breast tumors obtained from supplementary materials in Alexandrov et al., 2019 \citep{Alexandrov2019}.
  }
  \label{NAR-sigfig}
 \end{figure}

@@

\batcave resembles an empirical Bayes approach, but is not really empirical Bayes.
In fact, variants are best described as sequential samples from a data generating distribution represented by the mutation profile.
Early mutations are classified with low error, while later mutations are classified with less certainty.
We show here that sequential updating of the dirichlet-multinomial posterior converges to the true data generating distribution using the set of early mutations.
Along with the assumption that tumor evolution is approximately neutral, this allows us to generate a tumor and site-specific that sharpens the classification and calibration of later mutations.
Thus, \batcave is best described as an algorithm trained on early mutations.

>>>>

\begin{figure}
  \begin{center}
  \includegraphics{figures/signatures_only.pdf}
  \end{center}
  \bkmcomment{tumor in F is probably normal breast plus APOBEC. Mention somewhere?}
  \caption{Tumor mutation profiles.
  In each panel, the y-axis is the proportion of total mutations in the tumor with each of the 96 substitution types shown on the x-axis.
  \textbf{(A)} The observed mutation profile of an acute myeloid leukemia used in our real data analysis \cite{Griffith2015}.
  \textbf{(B)} The observed mutation profile of a breast tumor we used in our real data analysis \cite{Shi2018}.
  \textbf{(C)} A mutation profile used for simulating tumors, made up of equal proportions of COSMIC mutation signatures 1, 7, \& 11.
  \textbf{(D)} Equal proportions of signatures 1, 4, \& 5.
  \textbf{(E)} Equal proportions of signatures 1, 3, \& 5.
  \textbf{(F)}--\textbf{(G)} The observed mutation profile of two real breast tumors obtained from supplementary materials in Alexandrov et al., 2019 \citep{Alexandrov2019}.
  }
  \label{NAR-sigfig}
 \end{figure}

@@

\batcave resembles an empirical Bayes approach, but is not really empirical Bayes.
In fact, variants are best described as sequential samples from a data generating distribution represented by the mutation profile.
Early mutations are classified with low error, while later mutations are classified with less certainty.
We show here that sequential updating of the dirichlet-multinomial posterior converges to the true data generating distribution using the set of early mutations.
Along with the assumption that tumor evolution is approximately neutral, this allows us to generate a tumor and site-specific prior that sharpens the classification and calibration of later mutations.
Thus, \batcave is best described as an algorithm trained on early mutations.
